---
title: "U.S. State Maps"
author: "Adrienn Nagy"
format:
  html:
    code-fold: true
    toc: true
---

```{r}
#| include: FALSE

library(tidyverse)
library(stringr)
library(maps)
library(sf)
library(leaflet)
library(htmltools)
library(poliscidata)
library(viridis)
```

# Introduction

This page presents choropleth maps of U.S. states using both a numeric and a categorical variable. Each variable is shown using a static map and an interactive map.

Data source: `poliscidata::states`

---

```{r}
# Prepare mapping data

states_polygon <- as_tibble(map_data("state")) |>
  select(region, group, order, lat, long)

states_sf <- 
  read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson") |>
  select(name, geometry)

polisci_data <- as_tibble(poliscidata::states) |>
  select(state, carfatal07, cook_index3)

# Clean keys

polisci_data <- polisci_data |>
  mutate(state = str_to_lower(state),
         state = str_squish(state))

states_polygon <- states_polygon |>
  mutate(region = str_to_lower(region),
         region = str_squish(region))

states_sf <- states_sf |>
  mutate(name = str_to_lower(name),
         name = str_squish(name))

# Merge

state_plotting_polygon <- states_polygon |>
  left_join(polisci_data, by = c("region" = "state"))

state_plotting_sf <- states_sf |>
  left_join(polisci_data, by = c("name" = "state")) |>
  filter(!(name %in% c("alaska", "hawaii", "puerto rico")))
```

# Numeric Variable: Car Fatalities

This map shows the number of car fatalities by state.

```{r}
#| echo: false

state_plotting_polygon |>
  ggplot(aes(x = long, y = lat, group = group)) + 
  geom_polygon(aes(fill = carfatal07), color = "black") + 
  labs(fill = "Car Fatalities",
       title = "Car Fatalities by State") +
  coord_map() + 
  theme_void() +  
  scale_fill_viridis()
```

## Interactive Map

```{r}
#| echo: false
#| message: false
#| warning: false

pal_num <- colorNumeric("viridis", domain = state_plotting_sf$carfatal07)

state_plotting_sf <- state_plotting_sf |>
  mutate(labels_num = str_c(name, ": ", carfatal07))

labels_num <- lapply(state_plotting_sf$labels_num, HTML)

leaflet(state_plotting_sf) |>
  setView(-96, 37.8, 4) |>
  addTiles() |>
  addPolygons(
    fillColor = ~pal_num(carfatal07),
    weight = 2,
    opacity = 1,
    color = "white",
    fillOpacity = 0.7,
    label = labels_num
  )
```

**Insight:** Higher car fatality rates appear in parts of the South and central U.S., while lower rates are more common in coastal regions.

---

# Categorical Variable: Political Leaning (Cook Index)

This map shows political leaning of each state.

```{r}
#| echo: false

state_plotting_polygon |>
  ggplot(aes(x = long, y = lat, group = group)) + 
  geom_polygon(aes(fill = cook_index3), color = "grey") + 
  labs(fill = "Cook Index",
       title = "Political Leaning by State") +
  coord_map() + 
  theme_void() +  
  scale_fill_manual(values = c("red", "yellow", "blue"))
```

## Interactive Map

```{r}
#| echo: false
#| message: false
#| warning: false

pal_cat <- colorFactor(c("red", "yellow", "blue"), domain = state_plotting_sf$cook_index3)

state_plotting_sf <- state_plotting_sf |>
  mutate(labels_cat = str_c(name, ": ", cook_index3))

labels_cat <- lapply(state_plotting_sf$labels_cat, HTML)

leaflet(state_plotting_sf) |>
  setView(-96, 37.8, 4) |>
  addTiles() |>
  addPolygons(
    fillColor = ~pal_cat(cook_index3),
    weight = 2,
    opacity = 1,
    color = "white",
    fillOpacity = 0.7,
    label = labels_cat
  )
```

**Insight:** Republican-leaning states are mostly located in the South and central U.S., while Democratic-leaning states are concentrated on the West Coast and Northeast.
